Fish Comparisons Simple Models

### 2021-12-13 Fish Data Analysis/comp_fish_simple_models.Rmd
Back before adding roup level features

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(car)
library(viridis)
library(boot)
```

So here we're going to do some simpler model selection things with the data using GLMs instead of GAMs

@@

fold_angle_0_360_to_0_180 <- function(x){abs(abs(x-180)-180)}
```

@@

comp_data <- comp_data %>% mutate(Flow = ifelse(Flow == "0", "Flow 0", "Flow 2")) %>%
                           mutate(Ablation = ifelse(Ablation == "N", "No Ablation", "Ablated")) %>%
                           mutate(Darkness = ifelse(Darkness == "N", "Light", "Dark")) %>%
                           mutate(Heading_Diff = abs(Heading_Diff)) %>%
                           filter(Distance <= 4) %>%
                           filter(Speed_Diff <= 6) %>% 
                           mutate(quarter_heading_diff = abs(abs(Heading_Diff-90)-90)) %>%
                           mutate(Is_Aligned = ifelse(Heading_Diff < 30, 1, 0)) %>%
                           mutate(Is_Reversed = ifelse(Heading_Diff > 150, 1, 0)) %>%
                           mutate(Angle = fold_angle_0_360_to_0_180(Angle))

sum_comp_data <- comp_data %>% mutate(X_Distance = round_any(X_Distance,0.25), 
                                      Y_Distance = round_any(abs(Y_Distance),0.25)) %>%
                               group_by(Flow,Ablation,Darkness,X_Distance,Y_Distance) %>%
                               summarise(Speed_Diff = mean(Speed_Diff),
                                         Heading_Diff = ang_mean(Heading_Diff),
                                         Sync = mean(Sync),
                                         Quarter_Heading_Diff = mean(quarter_heading_diff),
                                         Is_Aligned = mean(Is_Aligned),
                                         Is_Reversed = mean(Is_Reversed))

@@

```{r}
speed_anova <- glm(Speed_Diff ~ Flow + Ablation + Darkness + Flow:Ablation + Flow:Darkness, data = comp_data)

heading_anova <- glm(quarter_heading_diff ~ Flow + Ablation + Darkness + Flow:Ablation + Flow:Darkness, data = comp_data)

sync_anova <- glm(Sync ~ Flow + Ablation + Darkness + Flow:Ablation + Flow:Darkness, data = comp_data)

dist_anova <- glm(Distance ~ Flow + Ablation + Darkness + Flow:Ablation + Flow:Darkness, data = comp_data)

```{r}
ggplot(comp_data, aes(x = Flow, y = Speed_Diff, fill = paste(Ablation,Darkness,sep="-")))+
  geom_boxplot(outlier.shape = NA) +
  ylim(0,3) +
  theme_light()

ggplot(comp_data, aes(x = Flow, y = quarter_heading_diff, fill = paste(Ablation,Darkness,sep="-")))+
  geom_boxplot(outlier.shape = NA) +
  ylim(0,90) +
  theme_light()

ggplot(comp_data, aes(x = Flow, y = Sync, fill = paste(Ablation,Darkness,sep="-")))+
  geom_boxplot(outlier.shape = NA) +
  ylim(0,1) +
  theme_light()

ggplot(comp_data, aes(x = Flow, y = Distance, fill = paste(Ablation,Darkness,sep="-")))+
  geom_boxplot(outlier.shape = NA) +
  ylim(0,4) +
  theme_light()
```

Now we add in distance and angle

```{r}
speed_anova_da <- glm(Speed_Diff ~ Distance*Angle*(Flow + Ablation + Darkness + Flow:Ablation + Flow:Darkness), data = comp_data)

Anova(speed_anova_da)

heading_anova_da <- glm(quarter_heading_diff ~ Distance*Angle*(Flow + Ablation + Darkness + Flow:Ablation + Flow:Darkness), data = comp_data)

Anova(heading_anova_da)

sync_anova_da <- glm(Sync ~ Distance*Angle*(Flow + Ablation + Darkness + Flow:Ablation + Flow:Darkness), data = comp_data)

Anova(sync_anova_da)
```

```{r}
ggplot(comp_data, aes(x = Distance, y = Speed_Diff, color = paste(Ablation,Darkness,sep="-"), fill = paste(Ablation,Darkness,sep="-")))+
  geom_smooth(method = "glm", formula = y ~ poly(x,4))+
  facet_wrap(~ Flow) +
  theme_light()

ggplot(comp_data, aes(x = Distance, y = quarter_heading_diff, color = paste(Ablation,Darkness,sep="-"), fill = paste(Ablation,Darkness,sep="-")))+
  geom_smooth(method = "glm", formula = y ~ poly(x,4))+
  facet_wrap(~ Flow) +
  theme_light()

ggplot(comp_data, aes(x = Distance, y = Sync, color = paste(Ablation,Darkness,sep="-"), fill = paste(Ablation,Darkness,sep="-")))+
  geom_smooth(method = "glm", formula = y ~ poly(x,4))+
  facet_wrap(~ Flow) +
  theme_light()
```
```{r}
ggplot(comp_data, aes(x = Angle, y = Speed_Diff, color = paste(Ablation,Darkness,sep="-"), fill = paste(Ablation,Darkness,sep="-")))+
  geom_smooth(method = "glm", formula = y ~ poly(x,4))+
  facet_wrap(~ Flow) +
  theme_light()

ggplot(comp_data, aes(x = Angle, y = quarter_heading_diff, color = paste(Ablation,Darkness,sep="-"), fill = paste(Ablation,Darkness,sep="-")))+
  geom_smooth(method = "glm", formula = y ~ poly(x,4))+
  facet_wrap(~ Flow) +
  theme_light()

ggplot(comp_data, aes(x = Angle, y = Sync, color = paste(Ablation,Darkness,sep="-"), fill = paste(Ablation,Darkness,sep="-")))+
  geom_smooth(method = "glm", formula = y ~ poly(x,4))+
  facet_wrap(~ Flow) +
  theme_light()
```

This creates a dataframe that is used for making model predictions and graphing them
```{r}
d <- seq(from = 0, to = 4, by = 0.1)
a <- seq(from = 0, to = 180, by = 10)

@@

predict_df_da <- expand.grid(Distance = d, Angle = a, Flow = flows, Ablation = ablation, Darkness = dark)
predict_df_da <- predict_df_da %>% mutate(X_Distance = Distance*(cos(deg2rad(Angle))), Y_Distance = Distance*(sin(deg2rad(Angle))))

predict_df_da <- predict_df_da %>% filter(!(Ablation == "Ablated" & Darkness == 'Dark'))
predict_df_da <- na.omit(predict_df_da)
```

### Speed Models

So I want to make a model that tries to include both distance and angle. What I really need is to figure out what power works best for angle and distance, and then include that. The issue is stepwise selection can just remove one of them entirely. So I'm going to use cross validation to select the power for both of them, and then use that. I will also calculate the AICs to help in this decision making process

```{r, warnings=F}
set.seed(7)

comp_data_speed_model <- comp_data %>% select(c(Flow,Darkness,Ablation,Angle,Distance,Speed_Diff))

max_poly <- 10

speed_dist_cv_error_10 <- rep(0,max_poly)
speed_dist_AIC_10 <- rep(0,max_poly)

for (i in 1:max_poly){
  speed_dist_fit <- glm(Speed_Diff ~ poly(Distance,i)*(Ablation+Flow+Darkness+Ablation:Flow+Darkness:Flow),
                        data = comp_data_speed_model)
  speed_dist_cv_error_10[i] <- cv.glm(comp_data_speed_model, speed_dist_fit, K = 10)$delta[1]
  speed_dist_AIC_10[i] <- AIC(speed_dist_fit)
}

speed_angle_cv_error_10 <- rep(0,max_poly)
speed_angle_AIC_10 <- rep(0,max_poly)

for (i in 1:max_poly){
  speed_angle_fit <- glm(Speed_Diff ~ poly(Angle,i)*(Ablation+Flow+Darkness+Ablation:Flow+Darkness:Flow),
                         data = comp_data_speed_model)
  speed_angle_cv_error_10[i] <- cv.glm(comp_data_speed_model, speed_angle_fit, K = 10)$delta[1]
  speed_angle_AIC_10[i] <- AIC(speed_angle_fit)
}

speed_poly_plot_df <- data.frame(c(seq(max_poly),seq(max_poly)),
                                 c(speed_dist_cv_error_10,speed_angle_cv_error_10),
                                 c(speed_dist_AIC_10,speed_angle_AIC_10),
                                 c(rep("Distance",max_poly),rep("Angle",max_poly)))

colnames(speed_poly_plot_df) <- c("Degree","Error","AIC","Predictor")

speed_poly_plot_df <- speed_poly_plot_df %>% group_by(Predictor) %>%
                                             mutate(minError = min(Error),minAIC = min(AIC)) %>%
                                             ungroup() %>%
                                             mutate(isMinEror = ifelse(Error == minError,3,1),isMinAIC = ifelse(AIC == minAIC,3,1))

ggplot(speed_poly_plot_df, aes(x = Degree, y = Error, color = Predictor))+
  geom_point(size = speed_poly_plot_df$isMinEror)+
  geom_line()+
  theme_light()+
  facet_wrap(~ Predictor, scales = "free") +
  scale_size(guide = "none")

ggplot(speed_poly_plot_df, aes(x = Degree, y = AIC, color = Predictor))+
  geom_point(size = speed_poly_plot_df$isMinAIC)+
  geom_line()+
  theme_light()+
  facet_wrap(~ Predictor, scales = "free") +
  scale_size(guide = "none")
```
Now let's use those polynomials to create the actual model

While they not be the lowest, the big improvements happen at 4 for Distance, and 2 for Angle

```{r}
speed_diff_glm <- glm(Speed_Diff ~ poly(Distance,4)*poly(Angle,2)*(Ablation+Flow+Darkness+Ablation:Flow+Darkness:Flow),
                        data = comp_data_speed_model)

summary(speed_diff_glm)
```

Now let's make some predictions

```{r}
speed_pred <- predict_df_da %>% mutate(Speed_Diff = predict(speed_diff_glm,predict_df_da))

comp_data <- comp_data %>% mutate(Round_Dist = as.factor(round_any(Distance,1)), Round_Angle = as.factor(round_any(Angle,30)))

ggplot()+
  geom_boxplot(data = comp_data, aes(x = Round_Dist, y = Speed_Diff))+
  facet_wrap(~ Flow + Ablation + Darkness) +
  theme_light()

round_dist_aov <- aov(Speed_Diff ~ Round_Dist*(Flow + Ablation + Darkness + Flow:Ablation + Flow:Darkness), data = comp_data)
Anova(round_dist_aov)

ggplot()+
  geom_boxplot(data = comp_data, aes(x = Round_Angle, y = Speed_Diff))+
  facet_wrap(~ Flow + Ablation + Darkness) +
  theme_light()

round_angle_aov <- aov(Speed_Diff ~ Round_Angle*(Flow + Ablation + Darkness + Flow:Ablation + Flow:Darkness), data = comp_data)
Anova(round_angle_aov)

# ggplot()+
#   geom_point(data = comp_data %>% filter(Speed_Diff <= 2), aes(x = Distance, y = Speed_Diff), alpha = 0.1)+
#   geom_smooth(data = speed_pred, aes(x = Distance, y = Speed_Diff))+
#   facet_wrap(~ Flow + Ablation + Darkness) +
#   theme_light()

# ggplot()+
#   geom_density_2d_filled(data = comp_data %>% filter(Speed_Diff <= 2), aes(x = Distance, y = Speed_Diff), contour_var = "ndensity")+
#   geom_smooth(data = speed_pred, aes(x = Distance, y = Speed_Diff, color = "red"))+
#   facet_wrap(~ Flow + Ablation + Darkness) +
#   theme_light()

# ggplot()+
#   geom_point(data = comp_data %>% filter(Speed_Diff <= 2), aes(x = Angle, y = Speed_Diff), alpha = 0.1)+
#   geom_smooth(data = speed_pred, aes(x = Angle, y = Speed_Diff))+
#   facet_wrap(~ Flow + Ablation + Darkness) +
#   theme_light()

# ggplot()+
#   geom_density_2d_filled(data = comp_data %>% filter(Speed_Diff <= 2), aes(x = Angle, y = Speed_Diff), contour_var = "ndensity")+
#   geom_smooth(data = speed_pred, aes(x = Angle, y = Speed_Diff, color = "red"))+
#   facet_wrap(~ Flow + Ablation + Darkness) +
#   theme_light()

ggplot(data = comp_data, aes(x = Angle, y = Distance, z = Speed_Diff))+
  stat_summary_2d() +
  facet_wrap(~ Flow + Ablation + Darkness) +
  scale_fill_viridis(direction = -1) +
  theme_light()
```
You know let's just try stepwise as well. There's jsut way too many thing in that possible model

```{r}

speed_m_all <- glm(Speed_Diff ~ Distance*Angle*(Ablation+Flow+Darkness+Ablation:Flow+Darkness:Flow) +
                                I(Distance^2)*I(Angle^2)*(Ablation+Flow+Darkness+Ablation:Flow+Darkness:Flow)+
                                I(Distance^3)*I(Angle^3)*(Ablation+Flow+Darkness+Ablation:Flow+Darkness:Flow)+
                                I(Distance^4)*I(Angle^4)*(Ablation+Flow+Darkness+Ablation:Flow+Darkness:Flow),
                        data = comp_data_speed_model)

speed_m_none <- glm(Speed_Diff ~ 1, data = comp_data_speed_model)

speed_m_both <- step(speed_m_none, direction = "both", scope = formula(speed_m_all), trace = F)

summary(speed_m_both)
```

### Sync Models

Now let's try the same thing for sync values (Rayleigh's R)

```{r, warnings=F}
set.seed(7)

comp_data_sync_model <- comp_data %>% select(c(Flow,Darkness,Ablation,Angle,Distance,Sync))

max_poly <- 10

sync_dist_cv_error_10 <- rep(0,max_poly)
sybc_dist_AIC_10 <- rep(0,max_poly)

for (i in 1:max_poly){
  sync_dist_fit <- glm(Sync ~ poly(Distance,i)*(Ablation+Flow+Darkness+Ablation:Flow+Darkness:Flow),
                        data = comp_data_sync_model)
  sync_dist_cv_error_10[i] <- cv.glm(comp_data_sync_model, sync_dist_fit, K = 10)$delta[1]
  sybc_dist_AIC_10[i] <- AIC(sync_dist_fit)
}

sync_angle_cv_error_10 <- rep(0,max_poly)
sync_angle_AIC_10 <- rep(0,max_poly)

for (i in 1:max_poly){
  sync_angle_fit <- glm(Sync ~ poly(Angle,i)*(Ablation+Flow+Darkness+Ablation:Flow+Darkness:Flow),
                         data = comp_data_sync_model)
  sync_angle_cv_error_10[i] <- cv.glm(comp_data_sync_model, sync_angle_fit, K = 10)$delta[1]
  sync_angle_AIC_10[i] <- AIC(sync_angle_fit)
}

sync_poly_plot_df <- data.frame(c(seq(max_poly),seq(max_poly)),
                                 c(speed_dist_cv_error_10,speed_angle_cv_error_10),
                                 c(speed_dist_AIC_10,speed_angle_AIC_10),
                                 c(rep("Distance",max_poly),rep("Angle",max_poly)))

colnames(sync_poly_plot_df) <- c("Degree","Error","AIC","Predictor")

sync_poly_plot_df <- sync_poly_plot_df %>% group_by(Predictor) %>%
                                             mutate(minError = min(Error),minAIC = min(AIC)) %>%
                                             ungroup() %>%
                                             mutate(isMinEror = ifelse(Error == minError,3,1),isMinAIC = ifelse(AIC == minAIC,3,1))

ggplot(sync_poly_plot_df, aes(x = Degree, y = Error, color = Predictor))+
  geom_point(size = sync_poly_plot_df$isMinEror)+
  geom_line()+
  theme_light()+
  facet_wrap(~ Predictor, scales = "free") +
  scale_size(guide = "none")

ggplot(sync_poly_plot_df, aes(x = Degree, y = AIC, color = Predictor))+
  geom_point(size = sync_poly_plot_df$isMinAIC)+
  geom_line()+
  theme_light()+
  facet_wrap(~ Predictor, scales = "free") +
  scale_size(guide = "none")
```
Now let's use those polynomials to create the actual model

While they not be the lowest, the big improvements happen at 2 for Distance, and 2 for Angle

```{r}
sync_diff_glm <- glm(Sync ~ I(Distance^2)*I(Angle^2)*(Ablation+Flow+Darkness+Ablation:Flow+Darkness:Flow),
                        data = comp_data_sync_model)

summary(sync_diff_glm)
```

Now let's make some predictions

```{r}
sync_pred <- predict_df_da %>% mutate(Sync = predict(sync_diff_glm,predict_df_da))

>>>>

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(car)
library(viridis)
library(boot)
library(mgcv)
library(forcats)
library(ggpubr)
```

Alright, so from all of this what do I know?

Speed Differences:
  Flow increases speed differences
  Ablation decreases speed differences
  Darkness increases speed differences
  Distance increases speed differences
  Angles around 90 degrees have a bigger difference in speed

Heading Differences:
  Fish are more aligned when they are around 90 degrees, slightly less in ablated fish

Synchronization:
  Darkness increases synchronization

NND:
  Darkness increases NND
  Flow decreases NND

Nearest Neighbor Bearing:
  In darkness fish are more in front or behind
  In flow fish are more in positioned alongside each other

@@

fold_angle_0_360_to_0_180 <- function(x){abs(abs(x-180)-180)}

fold_angle_neg_180_180_to_neg_90_90 <- function(x){ifelse(x > 90,90-abs(90-x),ifelse(x < -90, -90+abs(-90-x), x))}

min_n <- function(x,n){sort(x)[1:n]}
max_n <- function(x,n){sort(x,decreasing = TRUE)[1:n]}
```

@@

comp_data <- comp_data %>% mutate(Flow = ifelse(Flow == "0", "Flow 0", "Flow 2")) %>%
                           mutate(Ablation = ifelse(Ablation == "N", "No Ablation", "Ablated")) %>%
                           mutate(Darkness = ifelse(Darkness == "N", "Light", "Dark")) %>%
                           filter(Distance <= 4) %>%
                           filter(abs(Speed_Diff) <= 6) %>% 
                           mutate(Speed_Diff = abs(Speed_Diff)) %>%
                           mutate(fold_heading_diff = abs(fold_angle_neg_180_180_to_neg_90_90(Heading_Diff))) %>%
                           mutate(Is_Aligned = ifelse(Heading_Diff < 30, 1, 0)) %>%
                           mutate(Is_Reversed = ifelse(Heading_Diff > 150, 1, 0)) %>%
                           separate(Fish, c("Fish1", "Fish2"),sep="x") %>%
                           mutate(Flow = factor(Flow), Ablation = factor(Ablation), Darkness = factor(Darkness)) %>%
                           mutate(Flow = fct_relevel(Flow, c("Flow 0","Flow 2"))) %>%
                           mutate(Ablation = fct_relevel(Ablation, c("No Ablation","Ablated"))) %>%
                           mutate(Darkness = fct_relevel(Darkness, c("Light","Dark"))) %>%
                           mutate(rad_Angle = Angle) %>%
                           mutate(sin_Angle = sin(rad_Angle), cos_Angle = cos(rad_Angle)) %>%
                           mutate(Flow_Ablation_Darkness = factor(paste(Flow,Ablation,Darkness,sep=", ")))

                           #mutate(Angle = fold_angle_0_360_to_0_180(Angle))

sum_comp_data <- comp_data %>% mutate(X_Distance = round_any(X_Distance,0.25), 
                                      Y_Distance = round_any(abs(Y_Distance),0.25)) %>%
                               group_by(Flow,Ablation,Darkness,X_Distance,Y_Distance) %>%
                               summarise(Speed_Diff = mean(Speed_Diff),
                                         Heading_Diff = ang_mean(Heading_Diff),
                                         Sync = mean(Sync),
                                         Fold_Heading_Diff = mean(fold_heading_diff),
                                         Is_Aligned = mean(Is_Aligned),
                                         Is_Reversed = mean(Is_Reversed))

comp_data_switch_fish <- comp_data %>% mutate(Fish3 = Fish1) %>%
                                       mutate(Fish1 = Fish2) %>%
                                       mutate(Fish2 = Fish3) %>%
                                       select(-c(Fish3))

comp_data_NND <- bind_rows(comp_data, comp_data_switch_fish) %>% group_by(Year,Month,Day,Trial,Fish1,Tailbeat.Num) %>%
                                                                 filter(Distance == min(Distance)) %>%
                                                                 ungroup()

comp_data_near_3 <- bind_rows(comp_data, comp_data_switch_fish) %>% group_by(Year,Month,Day,Trial,Fish1,Tailbeat.Num) %>%
                                                                 slice_min(Distance, n = 3) %>%
                                                                 ungroup() %>%
                                                                 distinct(Distance, .keep_all= TRUE)
```

```{r}
speed_anova <- aov(Speed_Diff ~ Flow + Ablation + Darkness + Flow:Ablation + Flow:Darkness, data = comp_data_near_3)

heading_anova <- aov(fold_heading_diff ~ Flow + Ablation + Darkness + Flow:Ablation + Flow:Darkness, data = comp_data_near_3)

sync_anova <- aov(Sync ~ Flow + Ablation + Darkness + Flow:Ablation + Flow:Darkness, data = comp_data_near_3)

dist_anova <- aov(Distance ~ Flow + Ablation + Darkness + Flow:Ablation + Flow:Darkness, data = comp_data_NND)

Anova(dist_anova)

angle_anova <- aov(Angle ~ Flow + Ablation + Darkness + Flow:Ablation + Flow:Darkness, data = comp_data_NND)

```{r}
ggplot(comp_data_near_3, aes(x = cos(deg2rad(Angle))))+
  geom_histogram()+
  facet_wrap(~ Flow + Ablation + Darkness)

ggplot(comp_data_near_3, aes(x = sin(deg2rad(Angle))))+
  geom_histogram()+
  facet_wrap(~ Flow + Ablation + Darkness)
```

```{r}
ggplot(comp_data_near_3, aes(x = Flow, y = Speed_Diff, fill = interaction(Ablation,Darkness,sep=", ")))+
  geom_boxplot(outlier.shape = NA) +
  guides(fill = guide_legend(title = "Condition")) +
  scale_fill_manual(values=c("#EEEEEE", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Available Senses on Speed Difference") +
  xlab("") +
  ylab("Speed Difference (BL/s)") +
  ylim(-0.1,2.5) +
  theme_light()

ggplot(comp_data_near_3, aes(x = Flow, y = fold_heading_diff, fill = interaction(Ablation,Darkness,sep=", ")))+
  geom_boxplot(outlier.shape = NA) +
  guides(fill = guide_legend(title = "Condition")) +
  scale_fill_manual(values=c("#EEEEEE", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Available Senses on Heading Difference") +
  xlab("") +
  ylab("Heading Difference (Degrees)") +
  theme_light()

ggplot(comp_data_near_3, aes(x = Flow, y = Sync, fill = interaction(Ablation,Darkness,sep=", ")))+
  geom_boxplot(outlier.shape = NA) +
  guides(fill = guide_legend(title = "Condition")) +
  scale_fill_manual(values=c("#EEEEEE", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Available Senses on Tailbeat Synchonization") +
  xlab("") +
  ylab("Syncronization") +
  theme_light()

ggplot(comp_data_near_3, aes(x = Flow, y = Distance, fill = interaction(Ablation,Darkness,sep=", ")))+
  geom_boxplot(outlier.shape = NA) +
  guides(fill = guide_legend(title = "Condition")) +
  scale_fill_manual(values=c("#EEEEEE", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Available Senses on Nearest Neighbor Distance (NND)") +
  xlab("") +
  ylab("NND (BL)") +
  theme_light()

ggplot(comp_data_near_3, aes(x = Flow, y = Angle, fill = interaction(Ablation,Darkness,sep=", ")))+
  geom_boxplot(outlier.shape = NA) +
  guides(fill = guide_legend(title = "Condition")) +
  scale_fill_manual(values=c("#EEEEEE", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Available Senses on Nearest Neighbor Bearing") +
  xlab("") +
  ylab("Nearest Neighbor Bearing (Degrees)") +
  scale_y_continuous(breaks = c(30,60,90,120,150)) +
  theme_light()

@@

Trying GAMs

```{r}
speed_gam <- gam(Speed_Diff ~ s(Distance,Angle,by=Flow_Ablation_Darkness),
                              data = comp_data_near_3)

summary(speed_gam)

heading_gam <- gam(fold_heading_diff ~ s(Distance,Angle,by=Flow_Ablation_Darkness),
                              data = comp_data_near_3)

summary(heading_gam)

sync_gam <- gam(Sync ~ s(Distance,Angle,by=Flow_Ablation_Darkness),
                              data = comp_data_near_3)

summary(sync_gam)
```

Now we make some predictions for this.
```{r}
d <- seq(from = 0, to = 3, by = 0.1)
a <- seq(from = 0, to = 180, by = 10)

@@

predict_df <- expand.grid(Distance = d, Angle = a, Flow = flows, Ablation = ablation, Darkness = dark)

predict_df <- predict_df %>% mutate(Flow_Ablation_Darkness = factor(paste(Flow,Ablation,Darkness,sep=", "))) %>% 
                            filter(!(Ablation == "Ablated" & Darkness == 'Dark'))

predict_df <- predict_df %>% mutate(pred_speed_diff = predict.gam(speed_gam,predict_df),
                                    pred_heading_diff = predict.gam(heading_gam,predict_df),
                                    pred_sync = predict.gam(sync_gam,predict_df))

predict_df <- predict_df %>% group_by(Distance,Flow,Ablation,Darkness) %>%
                             mutate(mean_Speed_Distance = mean(pred_speed_diff), sd_Speed_Distance = sd(pred_speed_diff),
                                    mean_Heading_Distance = mean(pred_heading_diff), sd_Heading_Distance = sd(pred_heading_diff),
                                    mean_Sync_Distance = mean(pred_sync), sd_Sync_Distance = sd(pred_sync)) %>%
                             ungroup() %>%
                             group_by(Angle,Flow,Ablation,Darkness) %>%
                             mutate(mean_Speed_Angle = mean(pred_speed_diff), sd_Speed_Angle = sd(pred_speed_diff),
                                    mean_Heading_Angle = mean(pred_heading_diff), sd_Heading_Angle = sd(pred_heading_diff),
                                    mean_Sync_Angle = mean(pred_sync), sd_Sync_Angle = sd(pred_sync)) %>%
                             ungroup()
```

And now we graph those?

```{r}
ggplot(predict_df, aes(x = Distance, y = mean_Speed_Distance, color = interaction(Ablation,Darkness,sep=", "),
                       fill = interaction(Ablation,Darkness,sep=" "))) +
  geom_line() +
  #geom_point(data=predict_df, aes(x = Distance, y = pred_speed_diff, alpha = 0.15)) +
  geom_ribbon(aes(ymin = mean_Speed_Distance-sd_Speed_Distance, ymax = mean_Speed_Distance+sd_Speed_Distance),
              linetype = 0,
              alpha = 0.1) +
  facet_wrap(~ Flow) +
  guides(fill = "none", color = guide_legend(title = "Condition")) +
  scale_color_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  scale_fill_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Distance and Senses on Speed Difference") +
  xlab("Distance (BL)") +
  ylab("Speed Difference (BL/s)") +
  ylim(0,1.5) +
  theme_light()

ggplot(predict_df, aes(x = Distance, y = mean_Heading_Distance, color = interaction(Ablation,Darkness,sep=", "),
                       fill = interaction(Ablation,Darkness,sep=", "))) +
  geom_line() +
  #geom_point(data=predict_df, aes(x = Distance, y = pred_heading_diff, alpha = 0.15)) +
  geom_ribbon(aes(ymin = mean_Heading_Distance-sd_Heading_Distance, ymax = mean_Heading_Distance+sd_Heading_Distance),
              linetype = 0,
              alpha = 0.1) +
  facet_wrap(~ Flow) +
  guides(fill = "none", color = guide_legend(title = "Condition")) +
  scale_color_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  scale_fill_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Distance and Senses on Heading Difference") +
  xlab("Distance (BL)") +
  ylab("Heading Difference (Degrees)") +
  ylim(0,50) +
  theme_light()

ggplot(predict_df, aes(x = Distance, y = mean_Sync_Distance, color = interaction(Ablation,Darkness,sep=", "),
                       fill = interaction(Ablation,Darkness,sep=", "))) +
  geom_line() +
  #geom_point(data=predict_df, aes(x = Distance, y = pred_sync, alpha = 0.15)) +
  geom_ribbon(aes(ymin = mean_Sync_Distance-sd_Sync_Distance, ymax = mean_Sync_Distance+sd_Sync_Distance),
              linetype = 0,
              alpha = 0.1) +
  facet_wrap(~ Flow) +
  guides(fill = "none", color = guide_legend(title = "Condition")) +
  scale_color_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  scale_fill_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Distance and Senses on Synchonization") +
  xlab("Distance (BL)") +
  ylab("Synchonization") +
  ylim(0,1)+
  theme_light()
```
```{r}
ggplot(predict_df, aes(x = Angle, y = mean_Speed_Angle, color = interaction(Ablation,Darkness,sep=", "),
                       fill = interaction(Ablation,Darkness,sep=", "))) +
  geom_line() +
  #geom_point(data=predict_df, aes(x = Angle, y = pred_speed_diff, alpha = 0.15)) +
  geom_ribbon(aes(ymin = mean_Speed_Angle-sd_Speed_Angle, ymax = mean_Speed_Angle+sd_Speed_Angle),
              linetype = 0,
              alpha = 0.1) +
  facet_wrap(~ Flow) +
  guides(fill = "none", color = guide_legend(title = "Condition")) +
  scale_color_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  scale_fill_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Bearing and Senses on Speed Difference") +
  xlab("Bearing (Degrees)") +
  ylab("Speed Difference (BL/s)") +
  ylim(0,1.5) +
  theme_light()

ggplot(predict_df, aes(x = Angle, y = mean_Heading_Angle, color = interaction(Ablation,Darkness,sep=", "),
                       fill = interaction(Ablation,Darkness,sep=", "))) +
  geom_line() +
  #geom_point(data=predict_df, aes(x = Angle, y = pred_heading_diff, alpha = 0.15)) +
  geom_ribbon(aes(ymin = mean_Heading_Angle-sd_Heading_Angle, ymax = mean_Heading_Angle+sd_Heading_Angle),
              linetype = 0,
              alpha = 0.1) +
  facet_wrap(~ Flow) +
  guides(fill = "none", color = guide_legend(title = "Condition")) +
  scale_color_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  scale_fill_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Bearing and Senses on Heading Difference") +
  xlab("Angle (Degrees)") +
  ylab("Heading Difference (Degrees)") +
  ylim(0,50) +
  theme_light()

ggplot(predict_df, aes(x = Angle, y = mean_Sync_Angle, color = interaction(Ablation,Darkness,sep=", "),
                       fill = interaction(Ablation,Darkness,sep=", "))) +
  geom_line() +
 #geom_point(data=predict_df, aes(x = Angle, y = pred_sync, alpha = 0.15)) +
  geom_ribbon(aes(ymin = mean_Sync_Angle-sd_Sync_Angle, ymax = mean_Sync_Angle+sd_Sync_Angle),
              linetype = 0,
              alpha = 0.1) +
  facet_wrap(~ Flow) +
  guides(fill = "none", color = guide_legend(title = "Condition")) +
  scale_color_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  scale_fill_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Bearing and Senses on Synchonization") +
  xlab("Bearing (Degrees)") +
  ylab("Synchonization") +
  ylim(0,1)+
  theme_light()
```
Density vs Things

```{r}
dist_bin_size <- 1
angle_bin_size <- 30

comp_data_freqs <- comp_data_near_3 %>% mutate(Dist_bin = round_any(Distance, dist_bin_size), 
                                               Angle_bin = round_any(Angle, angle_bin_size)) %>%
                                        group_by(Flow,Darkness,Ablation, Dist_bin, Angle_bin,Flow_Ablation_Darkness) %>%
                                        mutate(Count = n()) %>%
                                        ungroup() %>%
                                        group_by(Flow,Darkness,Ablation,Flow_Ablation_Darkness) %>%
                                        mutate(Freq = Count/n()) %>%
                                        ungroup() %>%
                                        group_by(Flow,Darkness,Ablation,Dist_bin,Angle_bin,Flow_Ablation_Darkness) %>%
                                        summarise(Speed_Diff = mean(Speed_Diff),
                                                  Fold_Heading_Diff = ang_mean(fold_heading_diff),
                                                  Sync = mean(Sync),
                                                  Freq = mean(Freq))

```{r}
freq_speed_glm <- glm(Speed_Diff ~ Freq*(Flow+Darkness+Ablation+Flow:Ablation+Flow:Darkness), data = comp_data_freqs)

summary(freq_speed_glm)

freq_heading_glm <- glm(Fold_Heading_Diff ~ Freq*(Flow+Darkness+Ablation+Flow:Ablation+Flow:Darkness), data = comp_data_freqs)

summary(freq_heading_glm)

freq_sync_glm <- glm(Sync ~ Freq*(Flow+Darkness+Ablation+Flow:Ablation+Flow:Darkness), data = comp_data_freqs)

summary(freq_sync_glm)

freq_speed_gam <- gam(Speed_Diff ~ s(Freq,by=Flow_Ablation_Darkness),
                              data = comp_data_freqs)

summary(freq_speed_gam)

freq_heading_gam <- gam(Fold_Heading_Diff ~ s(Freq,by=Flow_Ablation_Darkness),
                              data = comp_data_freqs)

summary(freq_heading_gam)

freq_sync_gam <- gam(Sync ~ s(Freq,by=Flow_Ablation_Darkness),
                              data = comp_data_freqs)

summary(freq_sync_gam)
```

Now we make some predictions for this.
```{r}
f <- seq(from = 0.005, to = 0.07, by = 0.001)

freq_predict_df <- expand.grid(Freq = f, Flow = flows, Ablation = ablation, Darkness = dark)

freq_predict_df <- freq_predict_df %>% mutate(Flow_Ablation_Darkness = factor(paste(Flow,Ablation,Darkness,sep=", "))) %>% 
                            filter(!(Ablation == "Ablated" & Darkness == 'Dark'))

freq_predict_df <- freq_predict_df %>% mutate(pred_speed_diff = predict.gam(freq_speed_gam,freq_predict_df),
                                    pred_heading_diff = predict.gam(freq_heading_gam,freq_predict_df),
                                    pred_sync = predict.gam(freq_sync_gam,freq_predict_df))
